---
title: "Airline Passenger Satisfication Dataset Cleaning & EDA"
date: "2022-09-19"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    number_sections: yes
    df_print: paged
    toc_float: yes
    self_contained: yes
    code_folding: hide
    theme: flatly
---
```{r setup, include=FALSE, warning=FALSE}
library(tidyverse)
library(naniar)
library(mice)
library(nnet)
library(rsq)
knitr::opts_chunk$set(echo = TRUE, fig.align = "center",warning=FALSE)
```

# Data cleaning
## Load data
```{r load data}
dftrain <- read.csv("train.csv")
dftest <- read.csv("test.csv")
all <- df <- rbind(dftrain,dftest)
dim(all)
```

```{r}
head(all)
#str(all)
```

## Missing values
### Explore missing data: the percentage of missing data
```{r}
miss_var_summary(all) %>% filter(pct_miss>0)
```

```{r Visualize missing data}
md.pattern(all, rotate.names = TRUE)
```

### Impute the NA values
For the percentage of missing data is very small, and avoid throwing away any valuable information. We replace missing values with the prediction using a simple linear regression, as Arrival.Delay.in.Minutes has a high positive relationship with Departure.Delay.in.Minutes.

```{r}
plot(all$Departure.Delay.in.Minutes,all$Arrival.Delay.in.Minutes, pch = 1)
lm1 = lm(Arrival.Delay.in.Minutes ~ Departure.Delay.in.Minutes, data = all)
rsq(lm1) # rsquared is large

all$Arrival.Delay.in.Minutes[is.na(all$Arrival.Delay.in.Minutes)] = predict(lm1, data.frame(Departure.Delay.in.Minutes = all$Departure.Delay.in.Minutes[is.na(all$Arrival.Delay.in.Minutes)]))
```


```{r}
sum(is.na(all))
dim(all)
```

## Identify Reponse Variable;Imbalance data?
```{r}
df.tbl <- tibble(df)
df.pos <- df.tbl %>% filter(satisfaction=="satisfied")
df.neg <- df.tbl %>% filter(satisfaction=="neutral or dissatisfied")

tar.prop <- data.frame(Per=round(c(nrow(df.pos)/nrow(df.tbl),nrow(df.neg)/nrow(df.tbl)),3),Class=c("Satisfied","Neutral or dissatisfied"))
label= paste(tar.prop$Class,scales::percent(tar.prop$Per,0.01))
pie(tar.prop$Per,labels = label,col = c("darkorange","deepskyblue3"), main = "Percentage of two target classes")
```

## Data types 
### Find all columns in characters
```{r}
chas = sapply(all, class)[sapply(all, class) == "character"]
names(chas)
```

### Transform and Dummy
```{r}
#Change columns in characters to nomial variables or ordinal variables and set dummy variables for nomial varaiables.
all[names(chas)[1:3]] = lapply(all[,names(chas)[1:3]], class.ind)
all$Class[all$Class == "Eco"] = 0
all$Class[all$Class == "Eco Plus"] = 1
all$Class[all$Class == "Business"] = 2
all$Class = as.numeric(all$Class)
#all$Class = factor(all$Class, levels = c("Eco","Eco Plus","Business"), ordered = TRUE)
```

## Outliers
### Outliers of numeric variables
```{r}
par(mfrow = c(1,2))

y<-factor(all$satisfaction)
plot(all$Age,y,main = "Age")
boxplot(all$Age,main = "Age")

plot(all$Flight.Distance,y,main = "Flight.Distance")
boxplot(all$Flight.Distance,main = "Flight.Distance")

plot(all$Departure.Delay.in.Minutes,y,main = "Departure.Delay.in.Minutes")
boxplot(all$Departure.Delay.in.Minutes,main = "Departure.Delay.in.Minutes")

plot(all$Arrival.Delay.in.Minutes,y,main = "Arrival.Delay.in.Minutes")
boxplot(all$Arrival.Delay.in.Minutes,main = "Arrival.Delay.in.Minutes")

```


### Dataset without outliers
```{r}
no_outliers_all = subset(all, all$Departure.Delay.in.Minutes < 1500 & all$Arrival.Delay.in.Minutes < 1500)
dim(no_outliers_all)
```

# Visualization
## Correlation
```{r, fig.width=7,fig.height=7}
cor_data = no_outliers_all[-1][-1]
cor_data$satisfaction = class.ind(cor_data$satisfaction)
library(corrplot)
corr = cor(cor_data)
corrplot(corr,type = "upper")
```

## Do histograms/density estimations show anything interesting
### Age
```{r}
means <- df.tbl %>%
  group_by(satisfaction) %>%
  summarise(mean = mean(Age))

ggplot(df.tbl,aes(x=Age,color=factor(satisfaction),fill=factor(satisfaction))) + 
  scale_color_manual(values = c("deepskyblue3","darkorange"),labels=c("Neutral or dissatisfied","Satisfied")) +
  scale_fill_manual(values = c("deepskyblue3","darkorange"),labels=c("Neutral or dissatisfied","Satisfied"))+
  geom_density(alpha=0.5) + 
  geom_vline(data = means, aes(xintercept = mean, color=factor(satisfaction)),linetype="dashed") +
  labs(title = "Distributon of Age by Satisfication Level",
       x = "Age",
       y = "Density",
       fill="Satisfication Level",
       color="Satisfication Level")+
  theme_minimal()
```

### Inflight Service
```{r}
means <- df.tbl %>%
  group_by(satisfaction) %>%
  summarise(mean = mean(Inflight.service))

ggplot(df.tbl,aes(x=Inflight.service,color=factor(satisfaction),fill=factor(satisfaction))) + 
  scale_color_manual(values = c("deepskyblue3","darkorange"),labels=c("Neutral or dissatisfied","Satisfied")) +
  scale_fill_manual(values = c("deepskyblue3","darkorange"),labels=c("Neutral or dissatisfied","Satisfied"))+
  geom_bar(alpha=0.65,pos="identity",width=1) + 
  geom_vline(data = means, aes(xintercept = mean, color=factor(satisfaction)),linetype="dashed") +
  labs(title = "Distributon of Inflight Service by Satisfication Level",
       x = "Inflight Service",
       y = "Density",
       fill="Satisfication Level",
       color="Satisfication Level")+
  scale_x_continuous(breaks=seq(1:5), labels = seq(1:5))+
  theme_minimal()
```

### Flight distance & Travel type
```{r violin plot}
library(ggunchained)

ggplot(df.tbl,aes(x=Type.of.Travel,y=Flight.Distance,fill=factor(satisfaction)))+
  geom_split_violin()+
  scale_fill_manual(values = c("deepskyblue3","darkorange"),labels=c("Neutral or dissatisfied","Satisfied"))+
  labs(title = "Distributon of Flight Distance by Satisfication Level and Travel Type",
       x = "Travel Type",
       y = "Flight Distance",
       fill="Satisfication Level")+  
  theme_minimal()
```

### Class
```{r, fig.width=5,fig.height=3}
ggplot(df.tbl, aes(Class, fill = factor(satisfaction))) +
  geom_bar() + 
  labs(title = "Satisfication Level by Class",
       x = "Class",
       y = "Count",
       fill="Satisfication Level") +
  scale_fill_manual(values = c("deepskyblue3","darkorange"),labels=c("Neutral or dissatisfied","Satisfied"))+
  theme_minimal()
```

## PCA
```{r}
dat.clean <- tibble(no_outliers_all)
```

```{r}
class <- factor(dat.clean[["satisfaction"]])
num.dat <- dat.clean %>% select(!satisfaction)
pca_res <- prcomp(num.dat,scale=TRUE,center = TRUE)

q1g.df <- data.frame(PC1 =pca_res$x[,1],
                        PC2 =pca_res$x[,2],
                        class=class)

ggplot(q1g.df,aes(PC1,PC2,color=class)) + 
  geom_point() + 
  theme_minimal()
```


